{\itshape Advanced Robotics} A simple path optimization method for motion planning

### 2015-10-20 paper.tex
Add results description, change images, re-fill table

% The following packages can be found on http:\\www.ctan.org
\usepackage{caption}
\usepackage{color}
\usepackage{graphicx} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{mathrsfs}
\DeclareMathOperator*{\argmin}{argmin}
\usepackage{algorithm}
\usepackage{algpseudocode}  %\usepackage[noend]{algpseudocode}
\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{proof}
\newtheorem{property}{Property}
\usepackage[numbers]{natbib}

\makeatletter
%\let\NAT@parse\undefined % allow hyperref to work with ieeeconf.cls for bib numbers
\makeatother
\usepackage{hyperref}
%\usepackage{cite}

% Math shortcuts :
\newcommand\real{\mathbb{R}}
\newcommand\p{\mathbf{p}}
\newcommand\pii{\mathbf{p_{i,i+1}}}
\newcommand\pij{\mathbf{p_{i,j}}}
\newcommand\gT{\tilde{g}^T}
\newcommand\g{\tilde{g}}
\newcommand\CS{\mathcal{C}}
\newcommand\dimCS{n_\mathbf{C}}
\newcommand\body{{\cal B}}
\newcommand\Sone{\mathbf{S}^1}
\newcommand\Sthree{\mathbf{S}^3}
\newcommand\conf{\mathbf{q}}
\newcommand\xx{\mathbf{x}}
\newcommand\cost{C}
\newcommand\weight{W}
\newcommand\translation{\mathbf{t}}
\newcommand\tcolli{t_{coll\ i}}
\newcommand\po{\mathbf{po}}

@@

\begin{abstract}
Most algorithms in probabilistic sampling-based path planning compute 
collision-free paths made of straight line segments lying in the configuration 
space. Due
to the randomness of sampling, the paths make detours that need to be optimized.
The contribution of this paper is to propose a gradient-based (GB) algorithm that
transforms a polygonal collision-free path into a shorter one, while both:
\begin{itemize}
\item requiring mainly collision checking, and few time-consuming obstacle 
distance computation,
\item constraining only part of the configuration variables that may cause a 
collision, and not the entire configurations, and
\item reducing parasite motions that are not useful for the problem resolution.
\end{itemize}
The algorithm is simple and requires few parameter tuning. Experimental
results include navigation and manipulation tasks, e.g. an robotic arm 
manipulating throw a window and a PR2 robot working in a kitchen environment, 
and comparisons with a random shortcut optimizer.

@@

The idea of our method is to find a good trade-off between
the simplicity of “blind” methods like shortcut algorithm,
and the complexity of distance based optimization techniques.
The method iteratively shortens the initial path with gradient-based information.
When a collision is detected at a given iteration, the method backtracks to the
latest valid iteration and inserts a one-dimensional constraint
between the objects detected in collision. Only collisions between objects are 
evaluated, therefore no pre-processing of the
robot or environment models is necessary to increase distance computation speed. 
Preserving the problem geometry also 
guarantees that a solution can still be found, e.g. for narrow passages as holes or 
grippers. The underlying optimization algorithm is a quadratic program.

@@

????? This paper is based on a similar work, which has been sent to ICRA, with a different type of collision constraints.

@@

@@

\begin{figure}[b]
	\centering
	\includegraphics[width=15cm]{contact-points.pdf}
	\caption{Bodies representation in collision-configuration (left) and 
	backtracked free-configuration (right). A contact point (i.e. any point in 
	the intersection of the bodies) $\po_c$ can be returned by the Flexible 
	Collision Library~\cite{fcl}, used to detect collisions. The red $(\perp)$ 
	line is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
	direction which is allowed by the constraint.}
	\label{contact-points}
\end{figure}

@@

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame ${\cal R}_0$ of each of the two 
bodies (Figure~\ref{contact-points} left). We compute the associated points $\po_1$ 
and $\po_2$ in the free-configuration $\xx_{i}(\tcolli)$, and the difference 
$\mathbf{u}=\po_2-\po_1$ which will be used as the orthogonal direction the 
bodies can not move along at the path parameter $\tcolli$ 
(Figure~\ref{contact-points} right).

@@

Let $M^{*}\in SE(3)$ denote the homogeneous transformation between frames of $
\body_1$  and $\body_2$ in configuration 
$\xx_i(\tcolli)$. Let $M\in SE(3)$ denote the same transformation in configuration 
$\xx (\tcolli)$ where $\xx$ is any path 
in $\CS^{wp}$. Let $R\in SO(3)$ and $\translation\in\real^3$ 
(resp. $\translation^{*}$) be the rotation 
and translation parts of $M$ (resp. $M^{*}$).

We denote by $F$ the path constraint defined by the mapping from $\CS^{wp}$ to $
\real^3$ defined by
\begin{equation}\label{eq:constraint}
F (\xx) =  \mathbf{u}^T R^T (\translation^{*} - \translation)
\end{equation}
where the product $R^T (\translation^{*} - \translation)$ is used to express the translations difference in the global frame ${\cal R}_0$.

Note that $F$ vanishes for any path $\xx$ such that in configuration
$\xx(\tcolli)$ the relative position of $\body_2$ w.r.t. $\body_1$ is equal to $
\translation^{*}$ or if the relative position $\translation^{*} - \translation$ is orthogonal to $\mathbf{u}$.

When a collision is detected at $\tcolli$ along path $\xx_{i+1}$, we wish to add
constraint
\begin{equation}\label{eq:non-linear constraint}
F (\xx) = 0
\end{equation}
However, as this constraint is non-linear, we linearize it around $\xx_i$:
$$
\frac{\partial F}{\partial \xx}(\xx_i)(\xx -\xx_i) = 0
$$
for any later iteration $\xx_j$, $j\geq i+1$. Both $F$ and $\frac{\partial F}{\partial \conf}$ are provided by our path planning framework~\cite{hpp} as differentiable functions.

@@

Let $q_{k\,i}$ denote the waypoint $k$ along path $x_i$.
The constraint on the position at $t_{coll\,i}$ is a function that depends only
on two waypoints $q_{k\,i}$, $q_{k+1\,i}$, with $0\leq k\leq n$. There exists 
$\beta\in[0,1]$ such that for any $j$:
$$
\xx_j (t_{coll\,i}) = \conf_{k\,j} + \beta (\conf_{k+1\,j} - \conf_{k\,j})
$$
The path Jacobian $J_F$ of constraint~(\ref{eq:constraint}) w.r.t. $x$ is thus built by block using the previous constraint jacobian $\frac{\partial F}{\partial \conf}$ and $\beta$. ????
%Therefore, the Jacobian $J_F$ of constraint~(\ref{eq:constraint}) with respect to 
%$x$ is
%$$
%J_F = \left(\begin{array}{cccccc}
%0 & \cdots & (1-\beta)\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \
%\beta\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \cdots & 0
%\end{array}
%\right)
%$$

@@

Each time a new constraint is added to the resolution, we stack its jacobian $J_F$ to the previous ones. This corresponds to \texttt{addCollisionConstraints} in 
Algorithm~\ref{algo:gradient}. Then we refer to~\cite{nocedal2006numerical} for solving constrained quadratic programs (QP) (function \texttt{computeIterate} in 
Algorithm~\ref{algo:gradient}).

@@

@@

In this part we describe the GB path-optimizer Algorithm~\ref{algo:gradient} 
according to the previous step definitions. The main difficulty here is to handle 
the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
be traveled through, and also to return a collision-free path. Typically, we 
chose $\alpha_{init} = 0.2$ to process small steps. $\alpha_{init} = 0.5$ is more 
encountered in the optimization literature, procuring larger steps but with more 
risks of collision.

\begin{algorithm}
\begin{algorithmic}%[1] % line numerotation
\Procedure{Optimize}{$\xx_0$}
	\State $\alpha \gets \alpha_{init}$
	\State $minReached \gets false$
	\While{($\textbf{not}(noCollision \textbf{ and } minReached)$)}
		\State $\p = \texttt{computeIterate}()$
		\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
		\State $\xx_1 \gets \xx_0 + \alpha\,\p$
		\If{(\textbf{not}(\texttt{validatePath($\xx_1$)}))}
			\State $noCollision \gets false$
			\If{$(\alpha \neq 1)$}
				\State \texttt{addCollisionConstraints}()
				\State $\alpha \gets 1$
			\Else
				\State $\alpha \gets \alpha_{init}$
			\EndIf
		\Else
			\State $\xx_0 \gets \xx_1$
			\State $noCollision \gets true$
		\EndIf
	\EndWhile
	\Return $\xx_0$
\EndProcedure
\end{algorithmic}
\caption{Gradient-based (GB) path-optimization.} \label{algo:gradient}
\end{algorithm}

Each time a collision-constraint is added, 
the solution of the current QP is tested (i.e. $\alpha = 1$). If this path is collision-free, the algorithm returns it as the solution. Otherwise, 
smaller steps are iteratively applied and tested toward the minimum, and constraints are added each time a collision is found.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=7.8cm]{contact_points6.png}
	%\noindent\rule{8cm}{0.8pt}
	\includegraphics[width=8cm]{contact_points2potential.png}
	\caption{Path-optimization results on 2D-ponctual robots, moving around 
	green obstacles. Initial paths are in red, optimized ones in black. Grey paths 
	represent intermediate iterations, red and black dots resp. $\po_1$ and $\po_2$ 
	points, red segments $\mathbf{u}$ and blue circles collision-free configurations 
	of the backtracked paths 
	where constraints have been computed.}
	\label{2D_long}
\end{figure}

@@

In the included video, we present multiple situations where the GB algorithm has 
been tested and compared to RS~\cite{randomShortcutHPP}
(Algorithm~\ref{algo:random-shortcut}) and PRS ~\cite{partialrandomShortcutHPP} 
adapted from~\cite{Geraerts04clearancebased}.
The termination condition of RS allows it to try 
shortening the path until 5 iterations of non-improvement are reached 
(corresponding to $maxNbFailures$ in Algorithm~\ref{algo:random-shortcut}).
For PRS, first a shortcut is tried on each DOF between $\conf_0$ and $\conf_{n+1}$. Then, for each DOF, configurations are shot as in Algorithm~\ref{algo:random-shortcut} and the steering-method returns a path made by an interpolation on only the current DOF and the previous subpath for other DOF. If this path is collision-free, it is replaced as in RS. This process is also stopped when 5 iterations of non-improvement are reached.

@@

% analysis UR2:
On the double-arm problem (see Figure~\ref{trajectories} top-right), one arm has to get around an obstacle while the 
other is relatively far from obstacles. As expected, the initial path given by 
RRT-connect activates both arms to solve the problem. Contrary to RS, 
GB manages to cancel the rotation of the free-arm while optimizing the 
first arm motion. Initial path length 5.17 has been decreased to 4.55 by RS and to 3.20 by GB.

\begin{figure}[t]
	\centering
	\includegraphics[width=6cm]{fiad_qinit.png}
	\includegraphics[width=6cm]{fiad_qfinal.png}\\
	\includegraphics[width=6cm]{fiad_real.png}
	\includegraphics[width=6cm]{fiad-trajectorires.png}

	\caption{(Bottom left) An industrial use-case example proposed by Philips for 	
	the 
	Factory-in-a-Day project~\cite{factory-day-video}. A similar environment have 
	been created (top) to 
	illustrate that our method can comply with an industrial problem, where 
	initial and final configurations are constrained in boxes. Trajectories results 
	are presented (bottom right).}
	\label{fiad}
\end{figure}

% analysis UR5:
For the following example with a 6-axis manipulator arm (see Figure~\ref{trajectories} bottom-left)
%(Figure~\ref{ur5})
in a  cluttered environment, path reduction only manages to reduce some but not 
all detours. This behavior can be explained by a path over-constraining due to the 
important proximity of the robot to the obstacles during the whole path, and 
because there are not so many DOF to play on to easily avoid collisions and 
return a small-cost solution.
We advance this explanation because on the two following high-DOF examples of the 
video (and Figure~\ref{pr2_final}), results are better 
in terms of path length and quality.

% analysis PR2 alone:
In the example presented Figure~\ref{pr2_final} and in the video, the mobile 
40-DOF PR2 simply has to cross its arms from 
the left arm up position to the right arm up one. Lengths of the initial path, the 
RS optimized path and the GB optimized path are 
respectively 14.5, 6.34 and 3.62. The RRT-connect planner 
returns detours and activates non-useful DOF such as the head, the torso lift 
and the translation on 
the ground. Once again, RS will hardly optimize the mobile base 
translation of the robot and other unnecessary DOF uses, unlike our 
optimized-path which mainly results in moving the arms as expected 
(Figure~\ref{pr2_final} right).
%The removal 
%of the translation is also a consequence of our cost ponderation as described in 
%the previous section.
% NON ???

% analysis PR2 in kitchen:
We obtain similar results on the PR2 performing a manipulation task 
in a kitchen environment (see the video). The robot has to move 
his hands from the top to the bottom of a table.
Our optimizer manages to reduce the initial length 15.8 from the visibility-PRM 
planner and improves the path quality 
just adding transformation constraints between the table and the 
robot's arms and grippers. Thus, the robot just slightly moves 
backward and uses its arm DOF to avoid the table, instead of 
processing a large motion to get away from the table. With GB, the path length is downed to 4.59, against 8.24 for RS.

Optimization computation time and path length averages are presented for 30 runs of 
the PR2-crossing-arms in the Table~\ref{tab:results}. In the same way, time 
computation averages for the PR2-in-kitchen example have been 
calculated: $59.5\,\text{s}$ (GB) and $76.2\,\text{s}$ (RS).
Thus our method presents 
similar computation times and path lengths for mobile manipulation tasks. However, 
the disappearance of the mobile base motion in the gradient-based optimized path is 
shown in the right column of Table~\ref{tab:results}. Thus the problem addressed 
Figure~\ref{decoupled_DOF_optimization} is solved considering this DOF.

>>>>

% The following packages can be found on http:\\www.ctan.org
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{color}
\usepackage{graphicx} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{mathrsfs}
\DeclareMathOperator*{\argmin}{argmin}
\usepackage{algorithm}
\usepackage{algpseudocode}  %\usepackage[noend]{algpseudocode}
\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{proof}
\newtheorem{property}{Property}
\usepackage[numbers]{natbib}
\algnewcommand\algorithmicinput{\textbf{input:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\algorithmicoutput{\textbf{output:}}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
\makeatletter
%\let\NAT@parse\undefined % allow hyperref to work with ieeeconf.cls for bib numbers
\makeatother
\usepackage{hyperref}
%\usepackage{cite}

% Math shortcuts :
\newcommand\real{\mathbb{R}}
\newcommand\p{\mathbf{p}}
\newcommand\pii{\mathbf{p_{i,i+1}}}
\newcommand\pij{\mathbf{p_{i,j}}}
\newcommand\gT{\tilde{g}^T}
\newcommand\g{\tilde{g}}
\newcommand\CS{\mathcal{C}}
\newcommand\dimCS{n_\mathbf{C}}
\newcommand\body{{\cal B}}
\newcommand\Sone{\mathbf{S}^1}
\newcommand\Sthree{\mathbf{S}^3}
\newcommand\conf{\mathbf{q}}
\newcommand\xx{\mathbf{x}} % \x already defined in the class
\newcommand\cost{C}
\newcommand\weight{W}
\newcommand\translation{\mathbf{t}}
\newcommand\tcolli{t_{coll\ i}}
\newcommand\po{\mathbf{po}}
\newcommand\Jf{{J_f}}
\newcommand\kernel{\mbox{Ker }}

@@

\begin{abstract}
Most algorithms in probabilistic sampling-based path planning compute 
collision-free paths made of straight line segments lying in the configuration 
space. Due
to the randomness of sampling, the paths make detours that need to be optimized.
The contribution of this paper is to propose a gradient-based (GB) algorithm that
transforms a polygonal collision-free path into a shorter one, while both:
\begin{itemize}
\item requiring only collision checking, and not any time-consuming obstacle distance computation,
\item constraining only part of the configuration variables that may cause a 
collision, and not the entire configurations, and
\item reducing parasite motions that are not useful for the problem resolution.
\end{itemize}
The algorithm is simple and requires few parameter tuning. Experimental
results include navigation and manipulation tasks, e.g. an robotic arm 
manipulating throw a window and a PR2 robot working in a kitchen environment, 
and comparisons with a random shortcut optimizer.

@@

The idea of our method is to find a good trade-off between
the simplicity of “blind” methods like shortcut algorithm,
and the complexity of distance based optimization techniques.
The method iteratively shortens the initial path with gradient-based information.
When a collision is detected at a given iteration, the method backtracks to the
latest valid iteration and inserts a one-dimensional constraint
between the objects detected in collision. Only collisions between objects are 
evaluated, therefore no pre-processing of the
robot or environment models is necessary to increase distance computation speed. 
Respecting the problem geometry also 
preserves that a solution can still be found, e.g. for narrow passages as holes or 
grippers. The method is also repeatable since no randomness is introduced. The underlying optimization algorithm is a quadratic program.

@@

This paper is based on a similar work sent to a conference, where a different type of collision constraints was studied.

@@

Let us assume that at iteration $i$, $k$ constraints have been inserted before the current interation. These constraints are stored as lines of a Jacobian matrix
$$
\Jf_{i} = \left(\begin{array}{c}L_1 \\ \vdots \\ L_k\end{array}\right)
$$
where iterate $\p_i$ are constrained to be in the kernel of $\Jf_{i}$
$$
\Jf_{i} \p_i = 0
$$

\subsection*{New constraint}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=15.8cm]{contact-points.pdf}
	\caption{Bodies representation in collision-configuration (left) and 
	backtracked free-configuration (right). A contact point (i.e. any point in 
	the intersection of the bodies) $\po_c$ can be returned by the Flexible 
	Collision Library~\cite{fcl}, used to detect collisions. The red $(\perp)$ 
	hyperplan is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
	direction which is allowed by the constraint.}
	\label{contact-points}
\end{figure}

@@

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame ${\cal R}_0$ 
(Figure~\ref{contact-points} left). We compute the respective positions $\po_1$ and 
$\po_2$ of $\po_c$ in 
the local frames of $\body_1$ and $\body_2$ (Figure~\ref{contact-points} right).

Let $M_1 \in SE(3)$ (resp. $M_2 \in SE(3)$) denote the homogeneous transformation 
between frames of $\body_1$ (resp. $\body_2$) and the global frame in configuration 
$\xx_i(\tcolli)$. Let $\mathbf{u}$ be the unit vector 
linking the contact points in configuration 
$\xx_{i}(t_{coll\ i})$ (the latest collision-free path), expressed in the local 
frame of $\body_1$:
$$
\mathbf{u} = \frac{M_{1}^{-1} (\xx_i) M_2 (\xx_i) \po_2 - \po_1}{\|M_{1}^{-1} 
(\xx_i) M_2 (\xx_i) \po_2 - \po_1\|}
$$

Notice that $\mathbf{u}$ is well defined since configuration 
$\xx_{i}(t_{coll\ i})$ is collision-free.
Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
$$
f (\xx) = \left(M_{1}^{-1} (\xx) M_2 (\xx) \po_2 - \po_1 | \mathbf{u}\right)
$$
The constraint is defined by
\begin {equation}\label{eq:new-constraint}
f(\xx) - f(\xx_{i}) = 0.
\end {equation}
As this constraint is non-linear, it is linearized around $\xx_{i}$.

Let $q_{k\,i}$ denote the waypoint $k$ along path $x_i$.
The constraint on the position at $t_{coll\,i}$ is a function that depends only
on two waypoints $q_{k\,i}$, $q_{k+1\,i}$, with $0\leq k\leq n$. There exists 
$\beta\in[0,1]$ such that for any $j$:
$$
\xx_j (t_{coll\,i}) = \conf_{k\,j} + \beta (\conf_{k+1\,j} - \conf_{k\,j})
$$
The path Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ is thus built by matrix 
blocks using the previous constraint Jacobian $\frac{\partial f}{\partial \conf}$ 
and $\beta$, corresponding to \texttt{computeCollisionConstraint} in 
Algorithm~\ref{algo:gradient}.
Then, a line is added in the contraint Jacobian matrix (function 
\texttt{addCollisionConstraint}), and the constraint 
Jacobian becomes:
\begin {align*}
\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{k+1}\end{array}\right)\ \mbox {with}\\
L_{k+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
\end{align*}

@@

%Let $M^{*}\in SE(3)$ denote the homogeneous transformation between frames of $
%\body_1$  and $\body_2$ in configuration 
%$\xx_i(\tcolli)$. Let $M\in SE(3)$ denote the same transformation in configuration 
%$\xx (\tcolli)$ where $\xx$ is any path 
%in $\CS^{wp}$. Let $R\in SO(3)$ and $\translation\in\real^3$ 
%(resp. $\translation^{*}$) be the rotation 
%and translation parts of $M$ (resp. $M^{*}$).

%We denote by $F$ the path constraint defined by the mapping from $\CS^{wp}$ to $
%\real^3$ defined by
%\begin{equation}\label{eq:constraint}
%F (\xx) =  \mathbf{u}^T R^T (\translation^{*} - \translation)
%\end{equation}
%where the product $R^T (\translation^{*} - \translation)$ is used to express the translations difference in the global frame ${\cal R}_0$.

@@

%Therefore, the Jacobian $J_F$ of constraint~(\ref{eq:constraint}) with respect to 
%$x$ is
%$$
%J_F = \left(\begin{array}{cccccc}
%0 & \cdots & (1-\beta)\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \
%\beta\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \cdots & 0
%\end{array}
%\right)
%$$

@@

Then we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
programs (QP) (function \texttt{computeIterate} in 
Algorithm~\ref{algo:gradient}).

@@

\subsection*{Convergence Analysis}
For iteratively constrained problem, the risk of falling in an infinite loop 
that keeps adding a redundant constraint exists.
In this section, we demonstrate that, given the constraint definition and the 
function \texttt{solveRedundantConstraint}, our 
algorithm will converge.
Under the previous notations, it is straight-forward that
\begin{equation}\label{eq:decreasing-kernel}
\kernel \Jf_{i+1} \subset \kernel \Jf_{i}
\end{equation}
Let us assume that
\begin{equation}\label{eq:asumption-1}
L_{k+1}\p_i \not= 0.
\end{equation}
We will elaborate later on this asumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then
$$
\kernel J_i \not= \kernel J_{i+1}.
$$
From~(\ref{eq:decreasing-kernel}), we deduce that
$$
\dim (\kernel J_{i+1}) < \dim (\kernel J_i).
$$
This results proves that each added constraint is linearly independant from the 
previous ones, and thus our algorithm terminates in a finite number of iterations.

\subsubsection*{About asumption~(\ref{eq:asumption-1})}%
Let us consider the following expression:
$$
\frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
$$
This expression represents the velocity of point $\po_2$ in the reference frame of Body~1 projected onto $\mathbf{u}$ along the motion of the robot defined by the following trajectory:
\begin{equation}\label{eq:trajectory}
\conf (t) = \xx_{i}(t_{coll\ i}) + t\, \alpha_i \p_{i}.
\end{equation}
The motion is defined by taking a constant abscissa $t_{coll\ i}$ and be moving the 
path along iterate $\p_i$. Note that configuration $\xx_{i+1}(t_{coll\ i})$ is 
reached for $t=1$.
This expression being equal to 0 requires either
\begin {enumerate}
\item that $\po_1$ and $\po_2$ come to contact at zero velocity, or
\item the velocity of $\po_2$ in the reference frame of Body~1 is orthogonal to 
$\mathbf{u}$ at contact ($t=1$).
\end {enumerate}
The first case is very unlikely and almost never encountered in real applications, 
except if the user builds such a case on purpose.
The second case may appear and can be noted by a rank loss of $\Jf_{i+1}$. In this 
case, we do not directly insert the new constraint. Instead, we process intermediate 
iterations until the redundancy vanishes, as explained in function 
\texttt{solveRedundantConstraint}.

\begin{algorithm}
\begin{algorithmic}%[1] % line numerotation
\INPUT{$\xx_{Free}$ latest collision-free path, $\xx_{Coll}$ current path with collisions, $\p$ current iteration step}
\Require constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$ built from $\xx_{Coll}$ would produce a rank loss in constraint Jacobian $\Jf$
\State $solved \gets false$
\State $minReached \gets false$
\While{($\textbf{not}(solved)$)}
	\State $\p = \texttt{computeIterate}()$
	\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
	\State $\xx \gets \xx_{Free} + \frac{\alpha}{2}\,\p$
	\If{(\texttt{validatePath}($\xx_1$))}
		\State $\xx_{Free} \gets \xx$
	\Else
		\State $\xx_{Coll} \gets \xx$
	\EndIf
	\State $\frac{\partial f}{\partial \xx} (\xx_{Free}) \gets$\texttt{computeCollisionConstraint}($\xx_{Coll}$, $\xx_{Free}$)
	\State $solved \gets$ \texttt{isFullRank}($\Jf$ , $\frac{\partial f}{\partial \xx} (\xx_{Free})$)
\EndWhile
\end{algorithmic}
\caption{\texttt{solveRedundantConstraint}()} \label{algo:solveRedundant}
\end{algorithm}

After producing an intermediate iteration $\xx_{i+2}$ where $\alpha_i$ was 
halved, the obtained path is tested for collisions. If it is collision-free, the 
constraint is now linearized around it. Otherwise, new contact points are taken 
on it, and the constraint is linearized around the latest free path.

Doing so, either the new constraint makes the dimension of the kernel decrease, or 
makes the configuration in collision ($\xx_{i+1}$ or $\xx_{i+2}$ depending on the 
case above) and the latest collision-free path ($\xx_{i}$ or $\xx_{i+2}$) become 
closer and closer. And the continuity of $f$ implies that
$$
\frac{\partial f}{\partial \xx}(\xx_i)\p_i \ \ \ \mbox{and}\ \ \ \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
$$
converge toward the norm of the velocity of point $\po_2$ in the reference frame of 
body~1 along trajectory~(\ref{eq:trajectory}) when $\po_2$ and $\po_1$ come to 
contact. We already handled this case previously, admitting that the velocity norm 
would not vanish. Therefore, is~\ref{eq:asumption-1} verified.

@@

In this part we describe the GB path-optimizer Algorithm~\ref{algo:gradient} 
according to the previous step definitions. The main difficulty here is to handle 
the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
be traveled through, and also to return a collision-free path. Typically, we 
chose $\alpha_{init} = 0.2$ or 0.1 to process small steps. $\alpha_{init} = 0.5$ is 
more encountered in the optimization literature, procuring larger steps but with 
more risks of collision.

\begin{algorithm}
\begin{algorithmic}%[1] % line numerotation
\INPUT{Path to optimize $\xx_0$}
\OUTPUT{Optimized collision-free path $\xx_0$}
\State $\alpha \gets \alpha_{init}$
\State $minReached \gets false$
\While{($\textbf{not}(noCollision \textbf{ and } minReached)$)}
	\State $\p = \texttt{computeIterate}()$
	\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
	\State $\xx_1 \gets \xx_0 + \alpha\,\p$
	\If{(\textbf{not}(\texttt{validatePath($\xx_1$)}))}
		\State $noCollision \gets false$
		\If{$(\alpha \neq 1)$}
			\State \texttt{computeCollisionConstraint}($\xx_1$, $\xx_0$)
			\State \texttt{solveRedundantConstraint}()
			\State \texttt{addCollisionConstraint}()
			\State $\alpha \gets 1$
		\Else
			\State $\alpha \gets \alpha_{init}$
		\EndIf
	\Else
		\State $\xx_0 \gets \xx_1$
		\State $noCollision \gets true$
	\EndIf
\EndWhile
\end{algorithmic}
\caption{Gradient-based (GB) algorithm for path-optimization.} \label{algo:gradient}
\end{algorithm}

Each time a collision-constraint is added, 
the solution of the current QP is tested (i.e. $\alpha = 1$). If this path is collision-free, the algorithm returns it as the solution. Otherwise, 
smaller steps are iteratively applied and tested toward the minimum, and constraints are added each time a collision is found with a small step (i.e. $\alpha \neq 1$).

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=7.8cm]{contact_points6.png}
	\includegraphics[width=8cm]{contact_points2potential.png}
	\caption{Path-optimization results on 2D-ponctual robots, moving around 
	green obstacles. Initial paths are in red, optimized ones in black. Grey paths 
	represent intermediate iterations, red and black dots resp. $\po_1$ and $\po_2$ 
	points, red segments $\mathbf{u}$ and blue circles collision-free 
	configurations 
	of the backtracked paths 
	where constraints have been computed.}
	\label{2D_long}
\end{figure}

@@

In the included video and Figures~\ref{pr2_final},~\ref{fiad} and
 ~\ref{fig:trajectories}, we present multiple situations where the GB algorithm has 
been tested and compared to RS~\cite{randomShortcutHPP}
(Algorithm~\ref{algo:random-shortcut}) and PRS ~\cite{partialrandomShortcutHPP} 
adapted from~\cite{Geraerts04clearancebased}.
The termination condition of RS allows it to try 
shortening the path until 5 iterations of non-improvement are reached 
(corresponding to $maxNbFailures$ in Algorithm~\ref{algo:random-shortcut}).
For our implementation of PRS, first a shortcut is tried on each DOF between 
$\conf_0$ and $\conf_{n+1}$. 
Then, for each DOF, configurations are shot as in 
Algorithm~\ref{algo:random-shortcut} and the steering-method returns a path made by 
an interpolation on only 
the current DOF and the previous subpath for other DOF. If this path is collision-
free, it is replaced as in RS. This process is also stopped when 5 iterations of 
non-improvement are reached.

@@

\vspace{0.4cm}
% freeflyer puzzle
Before entering the manipulators examples, the GB algorithm is tested and compared 
on a freeflyer puzzle example (see Figure~\ref{fig:trajectories} top right), very 
encountered in the motion planning literature. The puzzle has to cross down the 
obstacle using the hole in the middle. The initial path planned with Visibility-PRM 
contains detours above and below the obstacle, as well as small motion in the hole. 
Lengths of initial, GB optimized, and RS optimized paths are 
respectively 23.4, 8.53 and 10.2. PRS fails to decrease the path length because in this problem, all DOF are closely related to cross the hole. This behaviour was also noticed and improve in the PRS implementation of~\cite{Geraerts04clearancebased}, acting on a group of DOF. The GB optimizer manages to modify the puzzle trajectory even in the hole, contrary to RS which simply copies the initial path when the trajectory gets near of the obstacle.

\vspace{0.4cm}

% double-arm:
On the double-arm problem (see Figure~\ref{fig:trajectories} top-left), one arm has 
to get around a cylinder obstacle while the 
other arm is relatively far from obstacles. As expected, the initial path given by 
RRT-connect activates both arms to solve the problem. Contrary to RS, 
GB manages to cancel the rotation of the free-arm while optimizing the 
first arm motion, even if the second arm is still moving because of the translation 
motion. The initial path length 5.17 is decreased to 4.55 by RS, to 4.06 by PRS and 
to 2.77 by GB using 3 collision-constraints with the cylinder obstacle.

\begin{figure}[t]
	\centering
	\includegraphics[width=5.8cm,height=4.7cm]{fiad_qinit.png}
	\includegraphics[width=5.8cm,height=4.7cm]{fiad_qfinal.png}\\
	\includegraphics[width=5.8cm,height=4.7cm]{fiad_real.png}
	\includegraphics[width=5.8cm,height=4.7cm]{fiad-trajectorires.png}
	\caption{(Bottom left) An industrial use-case example proposed by Philips for 	
	the Factory-in-a-Day project~\cite{factory-day-video}. A similar environment 	
	has been created (top) to illustrate that our method can comply with an 
	industrial problem, where initial and final configurations of the UR5 are 
	constrained in boxes. Trajectories results 
	are presented (bottom right): initial (RRT) path in red, RS optimization in 
	blue and GB optimization in green.}
	\label{fiad}
\end{figure}

% UR5-spheres:
Some problems were 
involving an a 6-axis manipulator arm, also called UR5, equipped with a bar or a 
gripper.
In a relatively free environment (Figure~\ref{fig:trajectories} middle right), 
results from GB and RS are similar (path length are respectively 2.03 and 1.60). 
Note also that the end-effector trajectory has completely changed from the initial 
one, the robot is easily passing between the spheres, keeping its end-effector 
above.

% UR5-fiad
For an UR5 working in a cluttered environment inspired from an industrial issue 
Figure~\ref{fiad}, path optimization efficiently returns a shorter solution, close 
to the result of RS and to what can be observed in reality~\cite{factory-day-video}. 
The initial path length 6.65 is decreased to 2.08 by RS, to 
1.98 by GB and unchanged by PRS.
